---
title: "Assesssment 1 - COVID-19 impact on digital learning"
author: "Student Name and ID"
date: "2023-03-10"
output:
  word_document: default
  html_document:
    df_print: paged
student_ID: 123
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Cleaning and Wrangling

**Describe and show the data cleaning and wrangling process' that you have completed. Provide justification on any transformation you have completed.**

<br> <br>

you can embed your code like this:

```{r, summary-stats, echo=TRUE}
x <- 1:10    # create an x variable
y <- 10:1    # create a y variable
dataf <- data.frame(x = x, y = y)

summary(dataf)
```

\newpage

## Data Visualisation

**Provide evidence of your visualisation techniques and the reason for your choice of visual tools.**

<br>

You can also embed plots, for example:

```{r, simple-plot, echo=FALSE}
x <- 1:10    # create an x variable
y <- 10:1    # create a y variable
dataf <- data.frame(x = x, y = y)
plot(dataf$x, dataf$y, xlab = "x axis", ylab = "y axis")
```

\newpage

## Findings

**Clearly explain your findings and provide a valid conclusion of your analysis.**

<!-- Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. -->
